---
title: "MaisLE - net based"
author: "Mirjam R. Rieger"
date: "`r Sys.Date()`"
output: html_document
---

<style>
.scroll5 {
    height: 500px;
    overflow-y: scroll;
}
</style>

```{r dataprep, echo = F, warning = F, message = F}

## model parameters
if(length(grep("zip", modn) == 1))   pars <- c("alpha", "beta", "gamma", "theta", 
                                               "z", "k", "etasq", "rhosq")
if(length(grep("zinb",  modn) == 1)) pars <- c("alpha", "beta", "gamma", "theta", 
                                               "z", "k", "etasq", "rhosq", "phi")
## specify data and variables for m1
if(type == "m1") {
  Spec.order <- SpecL
  dat.mod    <- dat[dat$species %in% Spec.order,]
  Spec.list  <- levels(as.factor(dat.mod$species))
}

## specify data and variables for m2
if(type == paste0("m2_", group)) {
  
  ## summarize data based on group
  tmp <- dat %>% group_by(across(all_of(c(group, "site_short", "netNr", "decade")))) %>%
    summarize(NInd_pd = sum(NInd_pd),
              dens_1000hm2 = sum(dens_1000hm2),
              NSpec_pg = n(), .groups = "drop")
  
  dat.mod <- as.data.frame(
    left_join(tmp, 
              unique(select(dat, -c("NInd_pd", "species", "species_sci", "species_eng", "dens_1000hm2", "Ind_tot", groupL))), 
              by = c("site_short", "netNr", "decade")))

  dat.mod$species <- as.factor(dat.mod[,group])
  Spec.list       <- Spec.order <- levels(as.factor(dat.mod$species))
}

## axes labels
if(LANG == "GER") {
  dens   <- "Fangdichte [Ind. pro 1000 hm²]"
  dens2  <- "Dichte"
  nI     <- "Anzahl Individuen"
  nS     <- "Anzahl Arten"
  dist   <- "Abstand zur Feldgrenze [m]"
  maize  <- paste0("Maisanteil in ", RAD/1000, " km [prop.]")
  woodS  <- paste0("Wald- & Gehölzanteil in ", RAD/1000, " km [prop.]")
  height <- "Pflanzenhöhe [cm]"
  weed   <- "Verunkrautung [prop.]"
  spec   <- "Arten"
  year   <- "Jahr"
  jday   <- "Fangdekade"
  site   <- "Probefläche"
}

if(LANG == "EN") {
  dens   <- "density [ind. per 1000 hm²]"
  dens2  <- "density"
  nI     <- "no. of individuals"
  nS     <- "no. of species"
  dist   <- "distance to field margin [m]"
  maize  <- paste0("prop. of maize within ", RAD/1000, " km [prop.]")
  woodS  <- paste0("prop. of wood within ", RAD/1000, " km [prop.]")
  height <- "plant height [cm]"
  weed   <- "weed infestation [prop.]"  
  spec   <- "species"
  year   <- "year"
  jday   <- "decade"
  site   <- "study site"
}

# ## test dataset
# set.seed(1376)
# dat.mod <- dat.mod %>% group_by(site_short) %>%
#   mutate(keep = ifelse(runif(1, 0, 1) > 0.6, T, F)) %>%
#   ungroup(); table(dat.mod$keep)
# 
# dat.mod <- dat.mod[dat.mod$keep,]

## transform data
dat.mod$dist_c    <- c(scale(round(dat.mod$distance_m, 0)))
#dat.mod$height_c  <- c(scale(dat.mod$Hoehe_cm))
dat.mod$maize_p   <- round(dat.mod[,paste0("maize_", RAD, "m")], 0)/100
dat.mod$woodS_p   <- round(dat.mod[,paste0("woodS_", RAD, "m")], 0)/100
dat.mod$weed_p    <- dat.mod$prop_weed/100
dat.mod$maize_c   <- c(scale(dat.mod$maize_p))
dat.mod$woodS_c   <- c(scale(dat.mod$woodS_p))
dat.mod$weed_c    <- c(scale(dat.mod$prop_weed))
dat.mod$jday      <- c(scale(dat.mod$julian_day))
dat.mod$effort100 <- dat.mod$effort/100
dat.mod$sp        <- as.numeric(as.factor(dat.mod$species))

## create distance matrix for sites
site.coor <- dat.mod %>%
  group_by(site_short) %>%
  summarize(lon = mean(lon),
            lat = mean(lat),
            .groups = "drop")

site.coor <- site.coor[order(site.coor$site_short),]

Dmat <- as.matrix(dist(site.coor[, c("lon", "lat")], diag = T, upper = T))


```

# Data exploration

## Sample size
```{r}
table(dat.mod$species)
```


## Proportion of zeros
```{r VpropZ, echo = F, warning = F, message = F}
## get prop. of zeros per species (group)
dat.mod %>% group_by(species) %>%
  summarise(propZ = round(length(NInd_pd[NInd_pd == 0]) / n(), 2))
```

  
## Distribution of response variable {.tabset}

### NInd
```{r Vhist, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
## histogram of response variable (with and without effort)
ggplot(dat.mod) +
  geom_histogram(aes(NInd_pd, fill = species)) +
  facet_wrap(~species, scales = "free") +
  xlab(nI) +
  scale_fill_viridis_d(direction = -1, end = 0.9, option = "viridis") +
  theme_bw(base_size = bs) +
  theme(legend.position = "none")
```

### density
```{r Vhist2, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
ggplot(dat.mod) +
  geom_histogram(aes(dens_1000hm2, fill = species)) +
  facet_wrap(~species, scales = "free") +
  xlab(dens) +
  scale_fill_viridis_d(spec, direction = -1, end = 0.9, option = "viridis") +
  theme_bw(base_size = bs) +
  theme(legend.position = "none")
```

### NInd (zero excl.)
```{r Vhist3, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
## histogram of response variable without zeros (with and without effort)
ggplot(dat.mod[dat.mod$NInd_pd != 0,]) +
  geom_histogram(aes(NInd_pd, fill = species)) +
  facet_wrap(~species, scales = "free") +
  xlab(nI) +
  scale_fill_viridis_d(spec, direction = -1, end = 0.9, option = "viridis") +
  theme_bw(base_size = bs) +
  theme(legend.position = "none")
```

### density (zero excl.)
```{r Vhist4, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
ggplot(dat.mod[dat.mod$NInd_pd != 0,]) +
  geom_histogram(aes(dens_1000hm2, fill = species)) +
  facet_wrap(~species, scales = "free") +
  xlab(dens) +
  scale_fill_viridis_d(spec, direction = -1, end = 0.9, option = "viridis") +
  theme_bw(base_size = bs) +
  theme(legend.position = "none")

```


## Explanatory variables {.tabset}

### distance
```{r Vdist, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot1(x = "distance_m", xlab = dist) + theme_light(base_size = bs)
```

  
### maize
```{r Vmaize, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot1(x = "maize_p", xlab = maize) + theme_light(base_size = bs)
```

  
### woodS
```{r VwoodS, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot1(x = "woodS_p", xlab = woodS) + theme_light(base_size = bs)
```

### plant height
```{r Vheight, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot1(x = "Hoehe_cm", xlab = height) + theme_light(base_size = bs)
```

### weed
```{r Vweed, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot1(x = "weed_p", xlab = weed) + theme_light(base_size = bs)
```
  
### jday
```{r Vjday, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot1(x = "julian_day", xlab = jday) + theme_light(base_size = bs)

# df.jday <- dat.mod %>% group_by(species, round(jday*2, 0)/2) %>% summarize(mean = mean(dens_1000hm2))
# df.jday <- df.jday %>% group_by(species) %>% mutate(max = max(mean))
# df.jday$top <- df.jday$mean == df.jday$max
# 
# ggplot(df.jday) +
#   geom_line(aes(x = `round(jday * 2, 0)/2`, y = mean),
#              pch = 1, size = 2) +
#   scale_color_viridis_d(year, direction = -1, end = 0.9) +
#   facet_wrap(~species, scales = "free") +
#   xlab(jday) +
#   ylab(dens) +
#   ylim(0, NA) +
#   theme_bw(base_size = bs)
# 
# df.jday <- df.jday[order(df.jday$species),]

```

### site
```{r Vsite, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
ggplot(dat.mod) +
  geom_point(aes(x = site_short, y = dens_1000hm2, color = as.factor(year)),
             pch = 1, size = 2) +
  scale_color_viridis_d(year, direction = -1, end = 0.9, option = "inferno", alpha = 0.5) +
  facet_wrap(~species, scales = "free_y") +
  xlab(site) +
  ylab(dens) +
  theme_bw(base_size = bs) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

### distance & maize
```{r Vdm, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "distance_m", xlab = dist, y = "maize_p", ylab = maize) + theme_light(base_size = bs)
```

### distance & woodS
```{r Vdw, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "distance_m", xlab = dist, y = "woodS_p", ylab = woodS) + theme_light(base_size = bs)
```
  
### distance & height  
```{r Vdh, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "distance_m", xlab = dist, y = "Hoehe_cm", ylab = height) + theme_light(base_size = bs)
```

### distance & weed  
```{r Vdwe, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "distance_m", xlab = dist, y = "weed_p", ylab = weed) + theme_light(base_size = bs)
```

### distance & jday 
```{r Vdj, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "distance_m", xlab = dist, y = "julian_day", ylab = jday) + theme_light(base_size = bs)
```

### maize & woodS
```{r Vmw, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "maize_p", xlab = maize, y = "woodS_p", ylab = woodS) + theme_light(base_size = bs)
```

### maize & height
```{r Vmh, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "maize_p", xlab = maize, y = "Hoehe_cm", ylab = height) + theme_light(base_size = bs)
```

### maize & weed
```{r Vmwe, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "maize_p", xlab = maize, y = "weed_p", ylab = weed) + theme_light(base_size = bs)
```

### maize & jday 
```{r Vmj, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "maize_p", xlab = maize, y = "julian_day", ylab = jday) + theme_light(base_size = bs)
```

### woodS & height
```{r Vwh, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "woodS_p", xlab = woodS, y = "Hoehe_cm", ylab = height) + theme_light(base_size = bs)
```

### woodS & weed
```{r Vwwe, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "woodS_p", xlab = woodS, y = "weed_p", ylab = weed) + theme_light(base_size = bs)
```

### woodS & jday 
```{r Vwj, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "woodS_p", xlab = woodS, y = "julian_day", ylab = jday) + theme_light(base_size = bs)
```

### height & weed  
```{r Vhwe, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "Hoehe_cm", xlab = height, y = "weed_p", ylab = weed) + theme_light(base_size = bs)
```

### height & jday 
```{r Vhj, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "Hoehe_cm", xlab = height, y = "julian_day", ylab = jday) + theme_light(base_size = bs)
```

### weed & jday 
```{r Vwej, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot2(x = "weed_p", xlab = weed, y = "julian_day", ylab = jday) + theme_light(base_size = bs)
```


# Model
The used model is `r modn`, the used landscape radius is `r RAD` m. Analysed species `r if(type == "m1") paste("are ", c(Spec.list))` `r if(type == paste0("m2_", group)) paste("group is ", group, " (", c(Spec.list), ")")`.    

```{r runmod, echo = F}

stan.data <- with(dat.mod, list(
    N = nrow(dat.mod),
    
    Nsp = length(unique(species)),
    Nyear = length(unique(year2)),
    Nsite = length(unique(site_short)),
    Nland = 4, ## change if necessary

    sp = sp,
    year = as.numeric(as.factor(year2)),
    site = as.numeric(as.factor(site_short)),
    jday = jday,

    land = matrix(c(dist_c,
                    maize_c,
                    woodS_c,
                    weed_c),
                  ncol = 4), ## change if necessary
    
    Dmat = Dmat,
    
    y = as.integer(NInd_pd),
    effort = effort100

))

## run model if RUN == TRUE
mod <- cmdstan_model(paste0("../stan/", modn, ".stan"))

if(RUN) {
  fit <- mod$sample(stan.data, chains = 4, parallel_chains = 4, refresh = 100)
  fit$save_object(file = paste0("../saved models/", type, "_", modn, "_", RAD, "m"))
}

## read model if RUN == FALSE
if(!RUN) {
  fit <- readRDS(paste0("../saved models/", type, "_", modn, "_", RAD, "m"))
}

# fit$cmdstan_diagnose()

## number of simulations
nsim <- prod(dim(fit$draws("etasq"))[1:2])
```

<div class="scroll5">
```{r ppsum, echo = F}
## summary per parameter
print(fit$summary(pars), n = nrow(fit$summary(pars))+10)

```
</div>

## Posterior predictive checks  {.tabset}

<div class="scroll5">
### areas
```{r ppc1, echo = F, fig.dim = c(9, 5)}
for(p in pars) {
  print(mcmc_areas(fit$draws(p),
                   prob = 0.80,
                   prob_outer = 0.95))
}
```
</div>

<div class="scroll5">
### dens_overlay
```{r ppc2, echo = F, fig.dim = c(9, 5)}
for(p in pars) {
  print(mcmc_dens_overlay(fit$draws(p)))
}
```
</div>

<div class="scroll5">
### trace
```{r ppc3, echo = F, fig.dim = c(9, 5)}
for(p in pars) {
  print(mcmc_trace(fit$draws(p)))
}
```
</div>


## posterior predictive stats   {.tabset}

<div class="scroll5">
### NInd
```{r pps1, echo = F, warning = F, message = F, fig.dim = c(4.5, 4)}
modsims <- fit$draws("y_sim", format = "df")

for(sp in 1:length(Spec.list)) {
  y    <- stan.data$y[which(stan.data$sp == sp)]

  yrep <- get_draws("y_sim")[, which(stan.data$sp == sp)]

  print(ppc_hist(y, yrep[runif(8, 1, nsim),]) + ggtitle(Spec.list[sp]))

  gmean  <- ppc_stat(y, yrep, stat = "mean") +
    ggtitle("mean") + theme(legend.position = "none")
  gsd    <- ppc_stat(y, yrep, stat = "sd") +
    ggtitle("sd") + theme(legend.position = "none")
  gpropZ <- ppc_stat(y, yrep, stat = "prop_zero") +
    ggtitle("propZ") + theme(legend.position = "none")
  gmax   <- ppc_stat(y, yrep, stat = "max") +
    ggtitle("max") + theme(legend.position = "none")

  grid.arrange(gmean, gsd, gpropZ, gmax, ncol = 2, nrow = 2)
}
```
</div>

<div class="scroll5">
### density
```{r pps2, echo = F, warning = F, message = F, fig.dim = c(4.5, 4)}

for(sp in 1:length(Spec.list)) {
  y    <- stan.data$y[which(stan.data$sp == sp)]/ stan.data$effort[which(stan.data$sp == sp)]
  yrep <- get_draws("y_sim")[, which(stan.data$sp == sp)]/ stan.data$effort[which(stan.data$sp == sp)][col(yrep)]

  print(ppc_hist(y, yrep[runif(8, 1, 100),]) + ggtitle(Spec.list[sp]))

  gmean  <- ppc_stat(y, yrep, stat = "mean") +
    ggtitle("mean") + theme(legend.position = "none")
  gsd    <- ppc_stat(y, yrep, stat = "sd") +
    ggtitle("sd") + theme(legend.position = "none")
  gpropZ <- ppc_stat(y, yrep, stat = "prop_zero") +
    ggtitle("propZ") + theme(legend.position = "none")
  gmax   <- ppc_stat(y, yrep, stat = "max") +
    ggtitle("max") + theme(legend.position = "none")

  grid.arrange(gmean, gsd, gpropZ, gmax, ncol = 2, nrow = 2)

}
```
</div>

## Posterior predictions {.tabset}

```{r pp, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
## number of simulations for graph
gsim <- round(runif(SIM, 1, nsim))

## get draws
modsims <- fit$draws(c("alpha", "beta", "gamma", "theta", "k"), format = "df")
write.csv(modsims, paste0("../output/Coefficient_estimates_", type, "_", modn, "_", RAD, "m.csv"), 
          row.names = F, fileEncoding = "latin1")

## get Bayes p-value (in %) for beta (~ 0% is decrease, ~ 100% is increase)
xx <- modsims[,grep("beta", colnames(modsims))]
for(i in 1:ncol(xx)) {
  x <- xx[,i]
  print(colnames(xx)[i])
  print(round((sum(x>0)/nrow(x) + 0.5*sum(x=0)/nrow(x))*100, 1))
}

## extract predictions
if(PREDICT) {
  

  pp <- data.frame() # for predictions
  ff <- data.frame() # for SIM simulations (to plot confidence)
  
  newdat <- list()
  
  ## decade
  newdat[["jday"]] <- with(stan.data, expand.grid(
      land1 = mean(land[,1]),
      land2 = mean(land[,2]),
      land3 = mean(land[,3]),
      land4 = mean(land[,4]),
      jday = seq(min(jday), max(jday), length = 100),
      effort = 10,
      site = 1))
  newdat[["jday"]]$jday2 <- newdat[["jday"]]$jday^2
  
  ## distance
  newdat[["dist"]] <- with(stan.data, expand.grid(
      land1 = seq(min(land[,1]), max(land[,1]), length = 100),
      land2 = mean(land[,2]),
      land3 = mean(land[,3]),
      land4 = mean(land[,4]),
      jday = seq(min(jday), max(jday), length = 5),
      effort = 10,
      site = 1))
  newdat[["dist"]]$jday2 <- newdat[["dist"]]$jday^2
  
  ## maize
  newdat[["maize"]] <- with(stan.data, expand.grid(
      land1 = mean(land[,1]),
      land2 = seq(min(land[,2]), max(land[,2]), length = 100),
      land3 = mean(land[,3]),
      land4 = mean(land[,4]),
      jday = seq(min(jday), max(jday), length = 5),
      effort = 10,
      site = 1))
  newdat[["maize"]]$jday2 <- newdat[["maize"]]$jday^2
  
  ## woodS
  newdat[["woodS"]] <- with(stan.data, expand.grid(
      land1 = mean(land[,1]),
      land2 = mean(land[,2]),
      land3 = seq(min(land[,3]), max(land[,3]), length = 100),
      land4 = mean(land[,4]),
      jday = seq(min(jday), max(jday), length = 5),
      effort = 10,
      site = 1))
  newdat[["woodS"]]$jday2 <- newdat[["woodS"]]$jday^2
  
  ## weed
  newdat[["weed"]] <- with(stan.data, expand.grid(
      land1 = mean(land[,1]),
      land2 = mean(land[,2]),
      land3 = mean(land[,3]),
      land4 = seq(min(land[,4]), max(land[,4]), length = 100),
      jday = seq(min(jday), max(jday), length = 5),
      effort = 10,
      site = 1))
  newdat[["weed"]]$jday2 <- newdat[["weed"]]$jday^2
  
  
  for(s in 1:max(stan.data$sp)) {
    
    dzi <- c(get_draws(paste0("theta[", s)))        # zi part
    dal <- rowMeans(get_draws(paste0("alpha[", s))) # average across years
    db1 <- c(get_draws(paste0("beta[", s, ",1]")))  # land1, distance
    db2 <- c(get_draws(paste0("beta[", s, ",2]")))  # land2, maize
    db3 <- c(get_draws(paste0("beta[", s, ",3]")))  # land3, woodS
    db4 <- c(get_draws(paste0("beta[", s, ",4]")))  # land4, weed
    dg1 <- c(get_draws(paste0("gamma[", s, ",1]"))) # jday
    dg2 <- c(get_draws(paste0("gamma[", s, ",2]"))) # jday2
    dk  <- rowMeans(get_draws(paste0("k[", s)))     # average across sites
    
    for(p in names(newdat)) {
      Xmat  <- model.matrix(~ land1 + land2 + land3 + land4 + 
                              jday + jday2 + 
                              log(effort) + site, data = newdat[[p]])
  
      XmatZ <- model.matrix(~ 1, data = newdat[[p]]) # could be excluded, since 1
      
      fitmat <- matrix(ncol = nsim, nrow = nrow(newdat[[p]]))
      
      for(j in 1:nsim) {
        
        # without plogis for zi (is already (0,1)), with exp for poisson
        fitmat[, j] <- (1 - XmatZ %*% dzi[j]) * # zi part (XmatZ could be excluded)
          exp(Xmat %*% c( # poisson part
            dal[j], # year
            db1[j], # land1 (distance)
            db2[j], # land2 (maize)
            db3[j], # land3 (woodS)
            db4[j], # land4 (weed)
            dg1[j],  # jday    
            dg2[j],  # jday2   
            1, # effort100
            dk[j] # site
            ))
      }
      
      tmp <- newdat[[p]]
      
      ## summarize across jday (for other predictors)
      if(p != "jday") {
        tmp <- cbind(tmp, fitmat)
        
        tmp <- aggregate(. ~land1 + land2 + land3 + land4 + effort + site, 
                            data = tmp, FUN = mean)
        
        fitmat <- as.matrix(tmp[,(ncol(tmp)-nsim+1):ncol(tmp)])
        dimnames(fitmat) <- NULL # remove dimnames otherwise it can not be merged with jday fitmat
        tmp <- tmp[,1:(ncol(tmp)-nsim)]
      
      }
      
      ## extract quantiles
      tmp$fit  <- apply(X = fitmat, MARGIN = 1, FUN = quantile, probs = 0.5)
      tmp$mean <- apply(X = fitmat, MARGIN = 1, FUN = mean)
      tmp$lwr  <- apply(X = fitmat, MARGIN = 1, FUN = quantile, probs = 0.025)
      tmp$upr  <- apply(X = fitmat, MARGIN = 1, FUN = quantile, probs = 0.975)
      
      ## add variables
      tmp$pred <- p
      tmp$sp <- s
      tmp$species <- Spec.list[s]
      
      ## merge with previous data
      pp <- rbind(pp, tmp)
      
      ## extract 100 simulations (for Confidence lines)
      df.sim <- as.data.frame(fitmat[,gsim])
      
      ## add variables
      df.sim$sp <- s
      df.sim$species <- Spec.list[s]
      df.sim$pred <- p
      
      ## merge with previous simulations
      ff <- rbind(ff, df.sim)
    
    }
    
    gc()
    
    write.csv(pp, paste0("../output/Predictions_", type, "_", modn, "_", RAD, "m.csv"), row.names = F, fileEncoding = "latin1")
    write.csv(ff, paste0("../output/Predictions_sim_", type, "_", modn, "_", RAD, "m.csv"), row.names = F, fileEncoding = "latin1")
  
  }
  
  pp$species <- factor(pp$species, levels = Spec.order)
  ff$species <- factor(ff$species, levels = Spec.order)
  dat.mod$species <- factor(dat.mod$species, levels = Spec.order)
  
  ## rescale parameters
  pp$julian_day <- rescale(var = dat.mod$julian_day, var_c = pp$jday)
  pp$distance_m <- rescale(var = dat.mod$distance_m, var_c = pp$land1)
  pp$maize_p    <- rescale(var = dat.mod$maize_p,    var_c = pp$land2)
  pp$woodS_p    <- rescale(var = dat.mod$woodS_p,    var_c = pp$land3)
  pp$weed_p     <- rescale(var = dat.mod$weed_p,     var_c = pp$land4)
  
  
  pp.ov <- pp %>% group_by(distance_m, maize_p, woodS_p, weed_p, julian_day, pred) %>%
    summarise(fit = sum(fit),
              mean = sum(mean),
              lwr = sum(lwr),
              upr = sum(upr),
              .groups = "drop")
  
  
  write.csv(pp, paste0("../output/Predictions_", type, "_", modn, "_", RAD, "m.csv"), 
            row.names = F, fileEncoding = "latin1")
  write.csv(ff, paste0("../output/Predictions_sim_", type, "_", modn, "_", RAD, "m.csv"), 
            row.names = F, fileEncoding = "latin1")
  write.csv(pp.ov, paste0("../output/Predictions_overall_", type, "_", modn, "_", RAD, "m.csv"),
  row.names = F, fileEncoding = "latin1")
  
}

pp <- read.csv(paste0("../output/Predictions_", type, "_", modn, "_", RAD, "m.csv"), 
               encoding = "latin1")
ff <- read.csv(paste0("../output/Predictions_sim_", type, "_", modn, "_", RAD, "m.csv"), 
               encoding = "latin1")
pp.ov <- read.csv(paste0("../output/Predictions_overall_", type, "_", modn, "_", RAD, "m.csv"),
encoding = "latin1")

pp$species <- factor(pp$species, levels = Spec.order)
ff$species <- factor(ff$species, levels = Spec.order)
dat.mod$species <- factor(dat.mod$species, levels = Spec.order)

```

### jday
```{r ppjday, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot3(x = "julian_day", xlab = jday, preds = pp[pp$pred == "jday",], sims = ff[ff$pred == "jday",])
plot3(x = "julian_day", xlab = jday, preds = pp[pp$pred == "jday",], sims = ff[ff$pred == "jday",], 
      plot.raw = F, plot.wrap = F)
plot3(x = "julian_day", xlab = jday, preds = pp[pp$pred == "jday",], sims = ff[ff$pred == "jday",], 
      plot.raw = F, plot.wrap = F, plot.sim = F)
#ggsave(paste0("../graphs/prefig/jday_", type, "_", RAD, "m.png"), width = 18, height = 12, units = "cm")
```

### distance
```{r ppdist, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot3(x = "distance_m", xlab = dist, preds = pp[pp$pred == "dist",], sims = ff[ff$pred == "dist",])
plot3(x = "distance_m", xlab = dist, preds = pp[pp$pred == "dist",], sims = ff[ff$pred == "dist",],
      plot.raw = F, plot.wrap = F)
plot3(x = "distance_m", xlab = dist, preds = pp[pp$pred == "dist",], sims = ff[ff$pred == "dist",],
      plot.raw = F, plot.wrap = F, plot.sim = F)
#ggsave(paste0("../graphs/prefig/dist_", type, "_", RAD, "m.png"), width = 18, height = 12, units = "cm")
```

### maize
```{r ppmaize, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot3(x = "maize_p", xlab = maize, preds = pp[pp$pred == "maize",], sims = ff[ff$pred == "maize",])
plot3(x = "maize_p", xlab = maize, preds = pp[pp$pred == "maize",], sims = ff[ff$pred == "maize",],
      plot.raw = F, plot.wrap = F)
plot3(x = "maize_p", xlab = maize, preds = pp[pp$pred == "maize",], sims = ff[ff$pred == "maize",],
      plot.raw = F, plot.wrap = F, plot.sim = F)
#ggsave(paste0("../graphs/prefig/maize_", type, "_", RAD, "m.png"), width = 18, height = 12, units = "cm")
```

### woodS
```{r ppwoodS, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot3(x = "woodS_p", xlab = woodS, preds = pp[pp$pred == "woodS",], sims = ff[ff$pred == "woodS",])
plot3(x = "woodS_p", xlab = woodS, preds = pp[pp$pred == "woodS",], sims = ff[ff$pred == "woodS",],
      plot.raw = F, plot.wrap = F)
plot3(x = "woodS_p", xlab = woodS, preds = pp[pp$pred == "woodS",], sims = ff[ff$pred == "woodS",],
      plot.raw = F, plot.wrap = F, plot.sim = F)
#ggsave(paste0("../graphs/prefig/woodS_", type, "_", RAD, "m.png"), width = 18, height = 12, units = "cm")
```

### weed
```{r ppweed, echo = F, warning = F, message = F, fig.dim = c(10, 5)}
plot3(x = "weed_p", xlab = weed, preds = pp[pp$pred == "weed",], sims = ff[ff$pred == "weed",])
plot3(x = "weed_p", xlab = weed, preds = pp[pp$pred == "weed",], sims = ff[ff$pred == "weed",],
      plot.raw = F, plot.wrap = F)
plot3(x = "weed_p", xlab = weed, preds = pp[pp$pred == "weed",], sims = ff[ff$pred == "weed",],
      plot.raw = F, plot.wrap = F, plot.sim = F)
#ggsave(paste0("../graphs/prefig/weed_", type, "_", RAD, "m.png"), width = 18, height = 12, units = "cm")
```

